\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-19 paper.qmd
Added some diagnostic plots

<!--To generate errors for different total number of bins-->
```{r}
#| label: errors-scurve
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
#buffer_vec <- seq(0.05, 0.1, by = 0.01)

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = bin1, 
                         y = MSE)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  geom_vline(xintercept = 11, linetype="solid",
             color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = 9, linetype=2,
             color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = 13, linetype=2,
             color = "#bdbdbd", linewidth=1) +
  scale_x_continuous(breaks = 5:19) +
  labs(x = expression(b[1]), y = "MSE") +
  ggtitle("(b)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))
```

>>>>

<!--To generate errors for different total number of bins-->

```{r}
#| label: bin-counts-scurve-with_buffer
## To initialize number of bins along the x-axis
bin1_vec_scurve <- c(7, 11, 17) #sqrt(NROW(training_data_scurve)/2)
#buffer_vec <- seq(0.05, 0.16, by = 0.01)
buffer_vec <- c(0.05, 0.07, 0.12)

centroids_df_scurve <- data.frame(matrix(nrow = 0, ncol = 0))

for (xbins in bin1_vec_scurve) {
  for (buff in buffer_vec) {

    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2

    scurve_model <- fit_highd_model(
      training_data = training_data_scurve,
      emb_df = umap_scurve_scaled,
      bin1 = xbins,
      r2 = r2,
      q = buff, 
      is_bin_centroid = TRUE,
      is_rm_lwd_hex = FALSE,
      col_start_highd = "x"
    )

    df_bin_centroids_scurve <- scurve_model$df_bin_centroids
    #df_bin_scurve <- scurve_model$df_bin

    ## Compute error
    df_bin_centroids_scurve <- df_bin_centroids_scurve |>
      mutate(bin1 = xbins,
             bin2 = bin2,
             b = bin1 * bin2,
             b_non_empty = NROW(df_bin_centroids_scurve),
             buff = buff)

    centroids_df_scurve <- bind_rows(centroids_df_scurve, df_bin_centroids_scurve)

b7_q_plt <- centroids_df_scurve |>
  filter(bin1 == 7) |>
  filter(buff == 0.12) |>
  ggplot(aes(x = bin_counts)) +
  geom_density() +
  xlab("bin count") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.ticks.x = element_line(linewidth = 0.5),
    axis.title.x = element_text(size = 7)

  )

b11_q_plt <- centroids_df_scurve |>
  filter(bin1 == 11) |>
  filter(buff == 0.07) |>
  ggplot(aes(x = bin_counts)) +
  geom_density() +
  xlab("bin count") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.ticks.x = element_line(linewidth = 0.5), 
    axis.title.x = element_text(size = 7)

  )

b17_q_plt <- centroids_df_scurve |>
  filter(bin1 == 17) |>
  filter(buff == 0.05) |>
  ggplot(aes(x = bin_counts)) +
  geom_density() +
  xlab("bin count") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.ticks.x = element_line(linewidth = 0.5),
    axis.title.x = element_text(size = 7)

  )
```

```{r}
#| label: fig-bin-counts-scurve-with_buffer
#| fig-cap: "Distribution of bin counts for differnt $b_1$ with specific $q$"
#| out-height: 30%

b7_q_plt + b11_q_plt + 
  b17_q_plt +
  plot_layout(ncol = 3)

```{r}
#| out-height: 30%

b7_q_plt2 <- centroids_df_scurve |>
  filter(bin1 == 7) |>
  ggplot(aes(x = std_counts,
             colour = as.factor(buff),
             group = as.factor(buff))) +
  geom_density() +
  scale_color_brewer(type = "qual") +
  xlab("standardised bin count") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.ticks.x = element_line(linewidth = 0.5),
    axis.title.x = element_text(size = 7)

  )

b11_q_plt2 <- centroids_df_scurve |>
  filter(bin1 == 11) |>
  ggplot(aes(x = std_counts,
             colour = as.factor(buff),
             group = as.factor(buff))) +
  geom_density() +
  scale_color_brewer(type = "qual") +
  xlab("standardised bin count") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.ticks.x = element_line(linewidth = 0.5),
    axis.title.x = element_text(size = 7)
  )

b17_q_plt2 <- centroids_df_scurve |>
  filter(bin1 == 17) |>
  ggplot(aes(x = std_counts,
             colour = as.factor(buff),
             group = as.factor(buff))) +
  geom_density() +
  scale_color_brewer(type = "qual") +
  xlab("standardised bin count") +
  theme(
    axis.text.x = element_text(size = 7),
    axis.ticks.x = element_line(linewidth = 0.5),
    axis.title.x = element_text(size = 7)
  )

b7_q_plt2 + b11_q_plt2 + b17_q_plt2
```

```{r}
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
#buffer_vec <- seq(0.05, 0.16, by = 0.01)
buffer_vec <- c(0.05, 0.07, 0.12)

error_df_scurve_q <- data.frame(matrix(nrow = 0, ncol = 0))

for (xbins in bin1_vec_scurve) {
  for (buff in buffer_vec) {

    bin2 <- calc_bins_y(bin1 = xbins, r2 = r2, q = buff)$bin2

    scurve_model <- fit_highd_model(
      training_data = training_data_scurve,
      emb_df = umap_scurve_scaled,
      bin1 = xbins,
      r2 = r2,
      q = buff, 
      is_bin_centroid = TRUE,
      is_rm_lwd_hex = FALSE,
      col_start_highd = "x"
    )

    ## Compute error
    error_df <- glance(
      df_bin_centroids = df_bin_centroids_scurve,
      df_bin = df_bin_scurve,
      training_data = training_data_scurve,
      newdata = NULL,
      type_NLDR = "UMAP",
      col_start = "x") |>
      mutate(bin1 = xbins,
             bin2 = bin2,
             b = bin1 * bin2,
             b_non_empty = NROW(df_bin_centroids_scurve),
             q = buff)

    error_df_scurve_q <- bind_rows(error_df_scurve_q, error_df)

ggplot(error_df_scurve_q, 
       aes(x = bin1, 
           y = MSE,
           colour = as.factor(q),
           group = as.factor(q))) + 
  geom_point(size = 1) +
  geom_line() +
  scale_color_brewer(type = "qual") +
  scale_x_continuous(breaks = 5:19) +
  labs(x = expression(b[1]), y = "MSE") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))

```{r}
#| label: errors-scurve
## To initialize number of bins along the x-axis
bin1_vec_scurve <- 5:19 #sqrt(NROW(training_data_scurve)/2)
#buffer_vec <- seq(0.05, 0.1, by = 0.01)

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = bin1, 
                         y = MSE)) + 
  geom_point(size = 1) +
  geom_line(linewidth = 0.05) + 
  geom_vline(xintercept = 11, linetype="solid",
             color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = 9, linetype=2,
             color = "#bdbdbd", linewidth=1) +
  geom_vline(xintercept = 13, linetype=2,
             color = "#bdbdbd", linewidth=1) +
  scale_x_continuous(breaks = 5:19) +
  labs(x = expression(b[1]), y = "MSE") +
  ggtitle("(b)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12, 
            angle=90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        plot.title = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(1, 'cm'),
        legend.key.width = unit(1, 'cm'),
        legend.key.size = unit(1, 'cm'))
```

### 2024-08-19 paper.tex
Added some diagnostic plots

@@

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-60-1.pdf}
\end{center}

>>>>

\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/fig-bin-counts-scurve-with_buffer-1.pdf}

\caption{\label{fig-bin-counts-scurve-with_buffer}Distribution of bin
counts for differnt \(b_1\) with specific \(q\)}

\begin{center}
\includegraphics[width=1\textwidth,height=0.3\textheight]{paper_files/figure-pdf/unnamed-chunk-34-1.pdf}
\end{center}

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-35-1.pdf}
\end{center}

@@

\begin{center}
\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-64-1.pdf}
\end{center}
